Cognate beginnings to bilingual lexical acquisition

### 2023-05-25 docs/index.qmd
Fix README

@@

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------

@@

```{r}
#| label: questionnaire
#| echo: false
#| message: false
#| warning: false
data("pool", package = "bvq")
n_categories <- distinct(pool, semantic_category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% 
    pull(n) %>%
    set_names(c("catalan", "spanish"))
```

@@

```{r}
#| label: fig-participants-time
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 10 
#| fig-cap: "Date at which responses were collected"
participants %>% 
    group_by(time_stamp) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1, colour = clrs[1]) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
    theme(legend.position = "top",
          legend.title = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 20, hjust = 1),
          axis.title.x = element_blank()) +

@@

```{r}
#| label: fig-participants-age
#| fig-width: 7
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = "grey60") +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(x = "Age (months)",
         y = "# participants") +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55),
                       breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r apatb-participants}
#| label: tab-participants
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan. Information about language degree of exposure (DoE) was provided by families before filling the questionnaire. In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish. A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish. A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants |>
    mutate_at(vars(starts_with("doe_")), 
              (\(x) floor(x * 10) / 10)) |> 
    mutate(
        age = cut(floor(age),
                  c(10, 14, 18, 22, 26, 30, 34, 36),
                  include.lowest = TRUE),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%",
                       "25-50%", 
                       "50-75%",
                       "75-100%"),
            include.lowest = TRUE
        ) |>
            as.character()
    ) |>
    count(age, doe_catalan) |>
    arrange(desc(doe_catalan)) |>
    pivot_wider(names_from = age, values_from = n) |>
    gt() |>
    sub_missing(everything(), everything(), missing_text = "--") |>
    cols_label(doe_catalan = "Catalan exposure") |>
    tab_spanner("Age (months)", 2:7) |>
    tab_style(cell_text(align = "left"),
              list(cells_body(doe_catalan),
                   cells_column_labels(doe_catalan)))
```

@@

```{r}
#| label: fig-items-freq
#| echo: false
#| message: false
#| warning: false
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Distribution of lexical frequency across items. Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations. Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency. The black dot and interval indicate the mean and standard deviation."
mean_sd <- function(x, ...) {
    m <- mean(x, ...)
    s <- sd(x, ...)
    data.frame(mean = m, ymin = m-s, ymax = m+s)
}

@@

```{r}
#| label: fig-items-nphon
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of number of phonemes across items"
items %>%
    count(language, n_phon) %>%
    ggplot(aes(x = n_phon, y = n)) +
    facet_wrap(~language, ncol = 1) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = n),
              position = position_nudge(y = 5),
              colour = "black") +
    labs(x = "Phonemes",
         y = "Number of words",
         fill = "Language",
         shape = "Language",
         colour = "Language") +
    scale_x_continuous(breaks = 1:15) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(legend.position = "none",
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          axis.ticks.x = element_line(),
          legend.title = element_blank())
```

@@

```{r}
#| label: fig-items-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 9
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n),
           labels = paste0(
               n, " (",
               scales::percent(prop, accuracy = 0.1), 
               ")")) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = labels),
              position = position_nudge(y = 3), 
              size = 3.5) +
    labs(x = "Levenshtein distance", y = "Number of words") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

@@

```{r}
#| label: fig-items-frequency-phonemes
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq) %>% 
    distinct(te, .keep_all = TRUE) %>% 
    count(n_phon, freq) %>% 
    ggplot(aes(n_phon, freq)) +
    geom_point(size = 2,
               alpha = 0.5,
               colour = "grey60", 
               shape = 1, 
               stroke = 1,
               position = position_jitter(width = 0.1)) +
    geom_smooth(method = "lm",
                formula = "y ~ x" ,
                colour = "black") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(freq, lv) %>% 
    ggplot(aes(freq, lv)) +
    geom_point(size = 2,
               shape = 1, 
               colour = "grey60",
               stroke = 1) +
    geom_smooth(method = "lm", 
                formula = "y ~ x", 
                colour ="black") +
    labs(y = "Levenshtein similarity", 
         x = "Lexical frequency (Zipf score)", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = scales::percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1,
               stroke = 1,
               colour = "grey60") +
    geom_smooth(method = "lm",
                formula = "y ~ x",
                size = 1,
                colour = "black") +
    labs(x = "# Phonemes", 
         y = "Levenshtein similarity", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank()) 
```

@@

```{r fig_prior-mean}
#| label: fig-prior-mean
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3
#| out-width: 90%
#| fig-cap: "Expected prior-predicted mean."
nd <- expand.grid(n_phon_std = 0,
                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
                  exposure_std = 0,
                  lv_std = 0)

@@

```{r}
#| label: fig-prior-predictions
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
    exposure_std = 0,
    lv_std = 0
)

@@

```{r tbl-coefs}
#| label: tbl-coefs
#| tbl-cap: "Posterior distribution of regression coefficients."
#| tbl-note: "Median: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
#| out-width: 80%
# summarise posterior draws
intercept_label <- "Intercepts (at 22 months)"

@@

## Marginal effects
```{r fig-marginal}
#| label: fig-marginal
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 4.5
#| out-width: 100%
#| eval: true
#| fig-cap: "Posterior marginal effects. Thin lines correspond to 50 mean predictions. Thicker lines indicate the median of those predictions. Different colours indicate different levels of cognateness (phonological similarity). Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure. Predictions for Comprehension are show on top and predictions for Comprehension and Production are shown on the bottom. In-sample predictions lie inside the grey rectangles."
clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")

@@

```{r}
#| label: fig-results-random-id-age
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_age <- ranef(model_fit)$id[,,"age_std"]

@@

```{r}
#| label: fig-results-random-id-nphon
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the number of phonemes by participant"
re_n_phon <- ranef(model_fit)$id[,,"n_phon_std"]

@@

```{r}
#| label: fig-results-random-id-exposure
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of lexical frequency by participant"
re_freq <- ranef(model_fit)$id[,,"exposure_std"]

@@

```{r}
#| label: fig-results-random-id-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of phonological similarity by participant"
re_lv <- ranef(model_fit)$id[,,"lv_std"]

@@

```{r}
# | label: fig-results-random-id-exposure-lv
# | echo: true
# | message: false
# | warning: false
# | fig-width: 7
# | fig-cap: "Posterior random slope of age by participant"
re_exposure_lv <- ranef(model_fit)$id[,,"exposure_std:lv_std"]

@@

```{r}
#| label: fig-results-random-id-age-doe
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
re_age_exposure <- ranef(model_fit)$id[,,"age_std:exposure_std"]

@@

```{r}
#| label: fig-results-random-id-age-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
re_age_lv <- ranef(model_fit)$id[,,"age_std:lv_std"]

@@

```{r}
#| label: fig-results-random-id-age-exposure-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
re_age_exposure_lv <- ranef(model_fit)$id[,,"age_std:exposure_std:lv_std"]

@@

```{r}
#| label: fig-results-random-te
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior probability of acquisition by translation equivalent"
re_te <- ranef(model_fit)$te[,,"Intercept"] %>% 
    as_tibble() %>% 
    rownames_to_column("te") %>% 
    janitor::clean_names() %>% 
    mutate(
        across(
            .cols = where(is.numeric),
            .fns = list(
                Comprehension = ~plogis(. + fixef(model_fit)[1]),
                Production = ~plogis(. + fixef(model_fit)[2])
            ), .names = "{.col}__{.fn}"
        )
    ) %>% 
    select(-c(estimate:q97_5)) %>% 
    pivot_longer(
        where(is.numeric), 
        names_to = c(".value", "type"),
        names_sep = "__"
    ) %>% 
    mutate(te = as.numeric(te))

@@

```{r}
#| label: fig-results-random-te-age
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by translation equivalent"
re_age <- ranef(model_fit)$te[,,"age_std"]

@@

```{r}
#| label: fig-results-random-te-nphon
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
re_n_phon <- ranef(model_fit)$te[,,"n_phon_std"]

@@

```{r}
#| label: fig-results-random-te-exposure
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_doe <- ranef(model_fit)$te[,,"exposure_std"]

@@

```{r}
#| label: fig-results-random-te-age-exposure
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_age_doe <- ranef(model_fit)$te[,,"age_std:exposure_std"]

@@

```{r}
#| label: fig-diagnostics-traceplots
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
gather_draws(model_fit, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
    mutate(.chain = paste("Chain ", .chain)) %>%
    ggplot(aes(.iteration, .value, colour = .chain)) +
    facet_wrap(~.variable, scales = "free_y") +
    annotate(geom = "rect", colour = NA,
             xmin = 0, xmax = dim(model_fit$fit)[1]/2,
             ymin = -Inf, ymax = Inf,
             alpha = 0.25) +
    geom_line() +
    labs(
        x = "Iteration",
        y = "Sample value",
        colour = "Chain"
    ) +
    scale_colour_manual(values = clrs[c(1, 3, 4, 5)]) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = 7)
    )
```

### Momdel convergence

```{r fig-rhats-neffs}
#| label: fig-rhats-neffs
#| fig-cap: "MCMC convergence diagnostic of all parameters in the model. A: distribution of the Gelman-Rubin (R-hat) scores. B: distribution of the ratio of effective sa"
#| echo: false
#| message: false
#| warning: false
#| fig-height: 3
#| fig-width: 6
#| out-width: 100%
model_convergence |> 
    ggplot(aes(.rhat)) +
    geom_histogram(binwidth = 0.001, colour = "white", fill = "grey60") +
    geom_vline(xintercept = 1.01, linetype = "dashed") +
    labs(x = "R-hat",
         y = "MCMC samples") +

@@

```{r}
sessioninfo::session_info()
```

>>>>

code-fold: true
warning: false

@@

```{r setup}
#| label: setup
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------

@@

```{r}
#| label: questionnaire
#| message: false
#| warning: false
data("pool", package = "bvq")
n_categories <- distinct(pool, semantic_category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% 
    pull(n) %>%
    set_names(c("catalan", "spanish"))
```

@@

```{r}
#| label: fig-participants-time
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 10 
#| fig-cap: "Date at which responses were collected"
participants %>% 
    group_by(time_stamp) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1, colour = clrs[1]) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
    theme(legend.position = "top",
          legend.title = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 20, hjust = 1),
          axis.title.x = element_blank()) +

@@

```{r}
#| label: fig-participants-age
#| fig-width: 7
#| fig-height: 3
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = "grey60") +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(x = "Age (months)",
         y = "# participants") +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55),
                       breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r fig-participants}
#| label: tab-participants
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan. Information about language degree of exposure (DoE) was provided by families before filling the questionnaire. In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish. A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish. A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants |>
    mutate_at(vars(starts_with("doe_")), 
              (\(x) floor(x * 10) / 10)) |> 
    mutate(
        age = cut(floor(age),
                  c(10, 14, 18, 22, 26, 30, 34, 36),
                  include.lowest = TRUE),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%",
                       "25-50%", 
                       "50-75%",
                       "75-100%"),
            include.lowest = TRUE
        ) |>
            as.character()
    ) |>
    count(age, doe_catalan) |>
    arrange(desc(doe_catalan)) |>
    pivot_wider(names_from = age, values_from = n) |>
    gt() |>
    sub_missing(everything(), everything(), missing_text = "--") |>
    cols_label(doe_catalan = "Catalan exposure") |>
    tab_spanner("Age (months)", 2:7) |>
    tab_style(cell_text(align = "left"),
              list(cells_body(doe_catalan),
                   cells_column_labels(doe_catalan)))
```

@@

```{r}
#| label: fig-items-freq
#| message: false
#| warning: false
#| fig-width: 6
#| fig-height: 3
#| fig-cap: "Distribution of lexical frequency across items. Lexical frequencies were extracted from the English CHILDES corpora, and mapped to their Catalan and Spanish translations. Therefore Catalan and Spanish word-forms that belong to the same translation equivalent had the same lexical frequency. The black dot and interval indicate the mean and standard deviation."
mean_sd <- function(x, ...) {
    m <- mean(x, ...)
    s <- sd(x, ...)
    data.frame(mean = m, ymin = m-s, ymax = m+s)
}

@@

```{r}
#| label: fig-items-nphon
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of number of phonemes across items"
items %>%
    count(language, n_phon) %>%
    ggplot(aes(x = n_phon, y = n)) +
    facet_wrap(~language, ncol = 1) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = n),
              position = position_nudge(y = 5),
              colour = "black") +
    labs(x = "Phonemes",
         y = "Number of words",
         fill = "Language",
         shape = "Language",
         colour = "Language") +
    scale_x_continuous(breaks = 1:15) +
    scale_fill_manual(values = clrs[c(1, 4)]) +
    theme(legend.position = "none",
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          axis.ticks.x = element_line(),
          legend.title = element_blank())
```

@@

```{r}
#| label: fig-items-lv
#| message: false
#| warning: false
#| fig-width: 9
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n),
           labels = paste0(
               n, " (",
               scales::percent(prop, accuracy = 0.1), 
               ")")) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = labels),
              position = position_nudge(y = 3), 
              size = 3.5) +
    labs(x = "Levenshtein distance", y = "Number of words") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

@@

```{r}
#| label: fig-items-frequency-phonemes
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 4
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq) %>% 
    distinct(te, .keep_all = TRUE) %>% 
    count(n_phon, freq) %>% 
    ggplot(aes(n_phon, freq)) +
    geom_point(size = 2,
               alpha = 0.5,
               colour = "grey60", 
               shape = 1, 
               stroke = 1,
               position = position_jitter(width = 0.1)) +
    geom_smooth(method = "lm",
                formula = "y ~ x" ,
                colour = "black") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-items-frequency-lv
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(freq, lv) %>% 
    ggplot(aes(freq, lv)) +
    geom_point(size = 2,
               shape = 1, 
               colour = "grey60",
               stroke = 1) +
    geom_smooth(method = "lm", 
                formula = "y ~ x", 
                colour ="black") +
    labs(y = "Levenshtein similarity", 
         x = "Lexical frequency (Zipf score)", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = scales::percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1,
               stroke = 1,
               colour = "grey60") +
    geom_smooth(method = "lm",
                formula = "y ~ x",
                size = 1,
                colour = "black") +
    labs(x = "# Phonemes", 
         y = "Levenshtein similarity", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank()) 
```

@@

```{r fig_prior-mean}
#| label: fig-prior-mean
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3
#| out-width: 90%
#| fig-cap: "Expected prior-predicted mean."
nd <- expand.grid(n_phon_std = 0,
                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
                  exposure_std = 0,
                  lv_std = 0)

@@

```{r}
#| label: fig-prior-predictions
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
    exposure_std = 0,
    lv_std = 0
)

@@

```{r tbl-coefs}
#| label: tbl-coefs
#| tbl-cap: "Posterior distribution of regression coefficients."
#| tbl-note: "Median: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| message: false
#| warning: false
#| out-width: 80%
# summarise posterior draws
intercept_label <- "Intercepts (at 22 months)"

@@

```{r fig-marginal}
#| label: fig-marginal
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 4.5
#| out-width: 100%
#| eval: true
#| fig-cap: "Posterior marginal effects. Thin lines correspond to 50 mean predictions. Thicker lines indicate the median of those predictions. Different colours indicate different levels of cognateness (phonological similarity). Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure. Predictions for Comprehension are show on top and predictions for Comprehension and Production are shown on the bottom. In-sample predictions lie inside the grey rectangles."
clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")

@@

```{r}
#| label: fig-results-random-id-age
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_age <- ranef(model_fit)$id[,,"age_std"]

@@

```{r}
#| label: fig-results-random-id-nphon
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the number of phonemes by participant"
re_n_phon <- ranef(model_fit)$id[,,"n_phon_std"]

@@

```{r}
#| label: fig-results-random-id-exposure
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of lexical frequency by participant"
re_freq <- ranef(model_fit)$id[,,"exposure_std"]

@@

```{r}
#| label: fig-results-random-id-lv
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of phonological similarity by participant"
re_lv <- ranef(model_fit)$id[,,"lv_std"]

@@

```{r}
# | label: fig-results-random-id-exposure-lv
# | message: false
# | warning: false
# | fig-width: 7
# | fig-cap: "Posterior random slope of age by participant"
re_exposure_lv <- ranef(model_fit)$id[,,"exposure_std:lv_std"]

@@

```{r}
#| label: fig-results-random-id-age-doe
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
re_age_exposure <- ranef(model_fit)$id[,,"age_std:exposure_std"]

@@

```{r}
#| label: fig-results-random-id-age-lv
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
re_age_lv <- ranef(model_fit)$id[,,"age_std:lv_std"]

@@

```{r}
#| label: fig-results-random-id-age-exposure-lv
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
re_age_exposure_lv <- ranef(model_fit)$id[,,"age_std:exposure_std:lv_std"]

@@

```{r}
#| label: fig-results-random-te
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior probability of acquisition by translation equivalent"
re_te <- ranef(model_fit)$te[,,"Intercept"] %>% 
    as_tibble() %>% 
    rownames_to_column("te") %>% 
    janitor::clean_names() %>% 
    mutate(
        across(
            .cols = where(is.numeric),
            .fns = list(
                Comprehension = ~plogis(. + fixef(model_fit)[1]),
                Production = ~plogis(. + fixef(model_fit)[2])
            ), .names = "{.col}__{.fn}"
        )
    ) %>% 
    select(-c(estimate:q97_5)) %>% 
    pivot_longer(
        where(is.numeric), 
        names_to = c(".value", "type"),
        names_sep = "__"
    ) %>% 
    mutate(te = as.numeric(te))

@@

```{r}
#| label: fig-results-random-te-age
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by translation equivalent"
re_age <- ranef(model_fit)$te[,,"age_std"]

@@

```{r}
#| label: fig-results-random-te-nphon
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
re_n_phon <- ranef(model_fit)$te[,,"n_phon_std"]

@@

```{r}
#| label: fig-results-random-te-exposure
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_doe <- ranef(model_fit)$te[,,"exposure_std"]

@@

```{r}
#| label: fig-results-random-te-age-exposure
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_age_doe <- ranef(model_fit)$te[,,"age_std:exposure_std"]

@@

```{r}
#| label: fig-diagnostics-traceplots
#| message: false
#| warning: false
#| fig-width: 12
gather_draws(model_fit, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
    mutate(.chain = paste("Chain ", .chain)) %>%
    ggplot(aes(.iteration, .value, colour = .chain)) +
    facet_wrap(~.variable, scales = "free_y") +
    annotate(geom = "rect", colour = NA,
             xmin = 0, xmax = dim(model_fit$fit)[1]/2,
             ymin = -Inf, ymax = Inf,
             alpha = 0.25) +
    geom_line() +
    labs(
        x = "Iteration",
        y = "Sample value",
        colour = "Chain"
    ) +
    scale_colour_manual(values = clrs[c(1, 3, 4, 5)]) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = 7)
    )
```

### Model convergence

```{r fig-rhats-neffs}
#| label: fig-rhats-neffs
#| fig-cap: "MCMC convergence diagnostic of all parameters in the model. A: distribution of the Gelman-Rubin (R-hat) scores. B: distribution of the ratio of effective sa"
#| message: false
#| warning: false
#| fig-height: 4
#| fig-width: 7
#| out-width: 100%
model_convergence |> 
    ggplot(aes(.rhat)) +
    geom_histogram(binwidth = 0.001, colour = "white", fill = "grey60") +
    geom_vline(xintercept = 1.01, linetype = "dashed") +
    labs(x = "R-hat",
         y = "MCMC samples") +

@@

```{r}
sessionInfo()
```
